---
title: 'Assignment 2: Data visualization'
author: "Tamas Nagy"
output: html_document
editor_options: 
  chunk_output_type: console
---

You will have to create 3 plots based on the datasets and instructions detailed below. You will find the plots themeselves in the `assignments/assignment_2_plots`. Your task is to write the code that will reproduce the plots as closely as possible.

# Skills needed to solve this assignment

-   Using R and RStudio, reading data
-   Reporting using RMarkdown
-   Using Git and Github (for submitting the task)
-   Data manipulation (e.g. dplyr, tidyr), and working with factors (forcats)
-   Data visuzlization (ggplot2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidytuesdayR)
```

## Task 1: Climbing expeditions

The 2020-09-22 TidyTueday datasets are about climbing expeditions. From the three datasets, use the "expeditions". Reproduce the plot below! Notice a few things:

-   Use `forcats::fct_lump()` to get the 15 most frequent peaks, and drop the "Other" category.
-   The bars are ordered by the sum of all expeditions (use `fct_reorder()`).
-   The bar colors use the viridis palette and light theme.

```{r}
with_palette(viridis)
library(viridisLite)
mydf = tt_load("2020-09-22")
peaks = mydf$peaks
expeditions = mydf$expeditions
expedition_count = expeditions %>% 
  mutate(peak_name = fct_lump_n(peak_name,15)) %>% 
  filter(peak_name != "Other") %>% 
  count(peak_name, season)


library(viridis)
myplot = expedition_count %>% 
  ggplot(aes(x=reorder(peak_name, n), n, fill=factor(season))) +
  geom_bar(stat = "identity") +
  coord_flip() + scale_fill_viridis(discrete = TRUE ) + theme_light()+
  labs(title = "The 15 most popular peaks stacked by season of expedition", y = "Number of expeditions", x = "", fill = "season") +
  theme(legend.position = "bottom")

myplot


```

## Task 2: PhDs awarded

The 2019-02-19 TidyTueday dataset is about phd-s awarded by year and field. There is only one dataset, it is called `phd_by_field`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all phd-s by broad fields.
-   To make the x axis breaks pretty, use `scales::pretty_breaks()`, to make the y axis labels comma formatted, use `scales::comma_format()`.
-   The line size is 1.2, the colors are from the brewer "Dark2" palette. The theme is set to minimal.

```{r}
library(tidytuesdayR)
library(scale)
phd_data = tt_load("2019-02-19")
phd_data = phd_data$phd_by_field

by_year = phd_data %>% 
  na.omit() %>% 
 group_by(broad_field, year) %>% 
  summarize(phd_sum = sum(n_phds)) 

phd_plot = by_year %>% 
  ggplot(aes(x = year, y = phd_sum, color = broad_field)) +
  geom_line(size=1.2) +
  scale_x_continuous( labels= c("2008","2010","2012","2014","2016"), breaks = c(2008,2010,2012,2014,2016),limits = c(2008,2017))+
  scale_y_continuous(label = comma) +
  theme_minimal() +
  ggtitle("Number of awarded Ph.D.-s in the US by year") +
  labs(color = "Broad field") +
  scale_color_brewer(palette = "Dark2") + 
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

phd_plot

```

## Task 3: Commute in the US

The 2019-11-05 TidyTueday dataset is about commuting to work in each city in the US by bike or on foot. There is only one dataset, it is called `commute`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all commutes by state.
-   Both axis scales are log transformed and the labels comma formatted, using `scales::comma_format()`
-   The point size is 2, . The theme is set to light.

```{r}
library(tidytuesdayR)
commute_data = tt_load("2019-11-05")$commute
commute_summary = commute_data %>% 
  group_by(state_abb,mode, state_region) %>% 
  summarize(total = sum(n)) %>% 
  pivot_wider(names_from = mode, values_from = total)

commute_plot = commute_summary %>% 
  ggplot(aes(x = Walk, y = Bike, color = state_region, label = state_abb))+
  geom_point(size = 2) +
  geom_text(aes(label=state_abb),show.legend = F, colour = "black", check_overlap = T) +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma, limits = c(10^1,10^6), breaks = c(10^2, 10^3, 10^4, 10^5)) +
  labs(title = "Number of people walking vs biking to work in each USA state", y = "Number of ppl walking to work (log N)", x = "Number of ppl biking to work (log N)", color = "State region") +
  theme_light()
  

commute_plot
```
